---
title: "Stealing behind tunnels"
format: pdf
editor: visual
---cita
---

## load packages

```{r}
library(tidyverse)
library(lme4)
library(multcomp)
library(ggplot2)
source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
source("./functions/glmmTMB_stability.r")
source("./functions/extract_ranef_gmmTMB.r")
source("./functions/boot_glmmTMB.r")
load("stealing_behind_tunnels_workspace.RData")
```

## load demographics

```{r}
demo.data <- read.csv(file = "data/counterbalancing_stealing_tunnels.csv") %>%
  mutate(Subject = fct_recode(as.factor(Subject), Bärli = "B\xe4rli")) %>%
  mutate(first_condition = fct_recode(as.factor(first_condition), control = "contr")) %>%
  mutate(Subject = str_replace_all(Subject, " ", ""))%>%
  mutate(first_condition = str_replace_all(first_condition, " ", ""))



length(levels(as.factor(demo.data$Subject)))
```

## load loopy data

```{r}
loopy.data <- read.csv("data/scoringproject_83_stealing_behind_tunnels_scoring.csv") %>%
  mutate(Subject = str_replace_all(Subject, " ", ""))%>%
  mutate(Subject = fct_recode(as.factor(Subject), Kira16 = "Kira")) %>%
  dplyr::select(Scoring:trial) %>%
  mutate(duration = Stop - Start)%>%
  full_join(demo.data) %>%
  mutate(condition = ifelse(((first_condition == "test" & trial <= 2)| (first_condition == "control" & trial > 2)), "test", "control")) %>%
  mutate(opaque_choice = as.numeric(ifelse(Value == "opaque", 1,
                                           ifelse(Value == "transparent", 0 , "")))) %>%
  mutate(trial_w_con = as.numeric(ifelse(trial>2, trial-2, trial)))

levels(as.factor(loopy.data$Subject))
levels(as.factor(loopy.data$first_condition))
table(loopy.data$Subject, loopy.data$condition)

```

```{r}
first_choice_data <- loopy.data %>%
  filter(Behaviour == "first_choice")


table(first_choice_data$Subject, first_choice_data$condition)
table(first_choice_data$Subject, first_choice_data$opaque_choice)
table(first_choice_data$Subject, first_choice_data$trial) #three missing trials --> dogs did not make a choice
#which subjects (rows) in the above table have missing trials?
xx<-table(first_choice_data$Subject, first_choice_data$trial)
sum(apply(xx<1, MARGIN=1, FUN=sum)) #2 subjects (each missing 1 test trial) + 1 trial in which the dog needed longer than 120 s to make a choice


length(levels(as.factor(first_choice_data$Subject)))
```

aggregate data

```{r}
first_choice_data.agg <- first_choice_data %>%
  group_by(Subject, condition,trial_w_con, first_condition) %>%
  summarise(mean_opaque_ind = mean(opaque_choice)) 

first_choice_data.agg2 <- first_choice_data.agg %>%
  group_by(condition, trial_w_con, first_condition) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))

first_choice_data.agg3 <- first_choice_data.agg %>%
  group_by(condition,  first_condition) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))
```


```{r}
first_choice_data.agg_pretest <- first_choice_data %>%
  group_by(condition, trial_w_con, firstchoice_pretest2) %>%
  summarise(mean_opaque = mean(opaque_choice, na.rm=TRUE), median_opaque = median(opaque_choice, , na.rm=TRUE), sd_opaque = sd(opaque_choice, na.rm=TRUE), se_opaque = sd(opaque_choice)/sqrt(length(opaque_choice)))

ggplot(data = first_choice_data.agg_pretest, aes(x = as.factor(trial_w_con), y= mean_opaque, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  ylim(0, 1) + 
  xlab("Trial within condition") + 
  ylab ("Mean proportion opaque choice")+ 
  facet_wrap(~firstchoice_pretest2)

```

### plot of first choice data

```{r}
ggplot(data = first_choice_data.agg2, aes(x = as.factor(trial_w_con), y= mean_opaque, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  ylim(0, 1) + 
  xlab("Trial within condition") + 
  ylab ("Mean proportion opaque choice")+ 
  facet_wrap(~first_condition)

```
```{r}
first_choice_data.agg.fp <- first_choice_data %>%
  group_by(condition) %>%
  summarise(mean_opaque_choice = mean(opaque_choice), se_opaque_choice = sd(opaque_choice)/sqrt(length(opaque_choice)))


plot_first_choice<-ggplot(data = first_choice_data.agg.fp, aes(x = condition, y= mean_opaque_choice))+
    geom_bar(stat = "identity", alpha=.8)+
  geom_errorbar(aes(ymin = mean_opaque_choice - se_opaque_choice, ymax = mean_opaque_choice + se_opaque_choice),width=0.4
                )+
  geom_hline(yintercept = 0.5, lty=2)+
  theme_classic() +
  ylim(0, 0.75) + 
  xlab("Condition") + 
  ylab("Probability to steal behind opaque first")
plot_first_choice
```
## remove dogs that always chose same side
```{r}
first_choice_data.agg4 <- first_choice_data %>%
  group_by(Subject) %>%
  summarise(mean_opaque_ind = mean(opaque_choice)) %>%
  filter(mean_opaque_ind>0 & mean_opaque_ind<1)

first_choice_data_wo_sidebias <- first_choice_data %>%
  inner_join(first_choice_data.agg4)

first_choice_data_wo_sidebias_agg <- first_choice_data_wo_sidebias %>%
  group_by(Subject, condition,trial_w_con) %>%
  summarise(mean_opaque_ind = mean(opaque_choice)) 

first_choice_data_wo_sidebias_agg2 <- first_choice_data_wo_sidebias_agg %>%
  group_by(condition, trial_w_con) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))
```

```{r}
ggplot(data = first_choice_data_wo_sidebias_agg2, aes(x = as.factor(trial_w_con), y= mean_opaque, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  ylim(0, 1) + 
  xlab("Trial within condition") + 
  ylab ("Mean proportion opaque choice")

```

## Behavioral data: first choice - condition comparison

```{r}
#scale and center covariate
first_choice_data$z.trial<-as.vector(scale(first_choice_data$trial_w_con, center = TRUE, scale=TRUE))
#center factor
first_choice_data$condition <- as.factor(first_choice_data$condition)
levels(first_choice_data$condition)
first_choice_data$condition.c=as.vector(scale(as.numeric(first_choice_data$condition==levels(as.factor(first_choice_data$condition))[2]), center=TRUE, scale= FALSE))

#remove trial in which dog needed more than 2 minutes to make a choice
str(first_choice_data)
first_choice_data<-first_choice_data %>% 
  filter(Subject!="Monty7"|trial!=1)

table(first_choice_data$trial,first_choice_data$Subject)
```

fit model

```{r}
mm1_choice=glmer(opaque_choice ~ condition+z.trial+first_condition+
                   (1+condition.c|Subject),
             data=first_choice_data, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mm1_choice)
drop1(mm1_choice, test="Chisq")
```

```{r}

mm1_choice_drop1 <- drop1(mm1_choice, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1)
mm1_choice_drop1


mm1_choice_aic<-drop1(mm1_choice, test="Chisq")[1, "AIC"]
mm1_choice_aic_drop1<-drop1(mm1_choice, test="Chisq")[, "AIC"][-1]
mm1_choice_aic_drop1_delta<- c(NA, round(mm1_choice_aic - mm1_choice_aic_drop1, 2))
mm1_choice_aic_drop1_delta_df <- data.frame(delta_aic = mm1_choice_aic_drop1_delta)
```

Check for collinearity

```{r}
library(car)
xx=lm(opaque_choice ~ condition+z.trial+first_condition, data=first_choice_data)
vif(xx)
```
#### confidence intervals 
```{r}
mm1_choice.ci=boot.glmm.pred(model.res=mm1_choice, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)
```

#### model stability
```{r}
m.stab.fc <- glmm.model.stab(model.res = mm1_choice)
m.stab.fc$detailed$warnings
as.data.frame(round(m.stab.fc$summary[1:4, -1], 2))

m.stab.plot(round(m.stab.fc$summary[, -1], 2))
```


#### output table - first choice

```{r}
mm1_choice_output_table <-
  bind_cols(as.data.frame(summary(mm1_choice)$coeff),
            mm1_choice_drop1,
            mm1_choice.ci$ci.estimates,
            m.stab.fc$summary[1:4, -1],
            mm1_choice_aic_drop1_delta_df) %>%
  dplyr::select(
    Estimate,
    SE = `Std. Error`,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    min,
    max,
    delta_aic
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:df), ~ format(round(.x, 2), nsmall = 2))) %>%
 mutate(across(.cols = c(LowerCI:max), ~ format(round(.x, 2), nsmall = 2))) %>%  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_choice_output_table, file = "saves/mm1_choice_output_table.csv")
```

```{r}
save.image("stealing_behind_tunnels_workspace.RData")
```

## comparison against chance

```{r}
first_choice_test_data <- first_choice_data %>%
  filter(condition == "test")

first_choice_control_data <- first_choice_data %>%
  filter(condition == "control")
```


```{r}
mm1_choice_test_io=glmer(opaque_choice ~ 1 +
                   (1|Subject),
             data=first_choice_test_data, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mm1_choice_test_io)
```
In the test condition, the dogs' performance did not deviate significantly from the chance value of 0.5 (z = -0.76, p = 0.45)
```{r}
mm1_choice_control_io=glmer(opaque_choice ~ 1 +
                   (1|Subject),
             data=first_choice_control_data, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mm1_choice_control_io)
```



## latency data

```{r}
latency_data <- loopy.data%>%
  mutate(Value = ifelse(Subject == "Sophi" & Start_Frame == 143869, "latency_to_2nd_choice", Value))%>% #correct scoring mistake
  filter(Behaviour == "latency_first_choice") %>%
  rename(latency = duration)

latency_data%>%filter(Subject=="Sophi")

latency_fc_data <- loopy.data%>%
  mutate(Value = ifelse(Subject == "Sophi" & Start_Frame == 143869, "latency_to_2nd_choice", Value))%>%
  filter(Behaviour == "latency_first_choice", Value == "latency") %>%
  rename(latency = duration)

latency_sc_data <- loopy.data%>%
  mutate(Value = ifelse(Subject == "Sophi" & Start_Frame == 143869, "latency_to_2nd_choice", Value))%>%
  filter(Behaviour == "latency_first_choice", Value == "latency_to_2nd_choice") %>%
  rename(latency = duration)

latency_fc_data%>%filter(Subject=="Sophi")

```

aggregate latency data
##Plot latency

```{r}
latency_data_agg <- latency_data %>%
  group_by(Subject, condition, Value) %>%
  summarise(mean_latency = mean(latency))

latency_data_agg2 <- latency_data_agg %>%
  group_by(condition, Value) %>%
  summarise(mean_latency2 = mean(mean_latency), se_latency = sd(mean_latency)/sqrt(length(mean_latency)))

latency_data_fp <- latency_data %>%
  filter(Value=="latency") %>% 
  group_by(Subject,condition) %>%
  summarise(mean_latency = mean(latency), se_latency = sd(latency)/sqrt(length(latency)))

latency_plot<-ggplot(latency_data_fp, aes(x = condition, y = mean_latency)) + 
  geom_boxplot()+
  geom_jitter(alpha=.5, width=.4)+
  #geom_line(aes(group = Subject), alpha=0.3, lty=2) +
  ylab("Latency to first choice (s)")+
  xlab("Condition")+
  geom_segment(aes(x = 1, y = 95, xend = 2, yend = 95))+
  geom_text(label="*", x=1.5, y=100,size=7)+
  theme_classic()
latency_plot


library(cowplot)
Fig5<-plot_grid(plot_first_choice, latency_plot, nrow=1, labels=c("A)", "B)"))
Fig5
```

```{r}
ggsave(Fig5, filename = "graphics/Fig5.png", width = 9, height=6, scale=0.6)
```


```{r}
ggplot(latency_data_agg, aes(x = condition, y = mean_latency)) + 
  geom_boxplot(outlier.colour = "white")+
  geom_jitter()+
  facet_wrap(~Value)+ 
  theme_bw()
```

```{r}
ggplot(latency_data_agg2, aes(x = condition, y = mean_latency2)) +
  geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_latency2 - se_latency, ymax = mean_latency2 + se_latency  ),  position=position_dodge(.9), width=0.4)+
  facet_wrap(~Value)+ 
  theme_bw()

```


#### Model preparation
```{r}
latency_fc_data<- latency_fc_data%>%
  dplyr::select(Subject, latency, condition, trial_w_con, first_condition)%>%
  full_join(first_choice_data%>%
  dplyr::select(Subject, opaque_choice, condition, trial_w_con, first_condition))%>%
  mutate(choice = fct_recode(as.factor(opaque_choice), "opaque" ="1", "clear" ="0"))

hist(latency_fc_data$latency)
max(latency_fc_data$latency)


table(latency_fc_data$Subject, latency_fc_data$choice)
xx<-table(latency_fc_data$Subject, latency_fc_data$trial)
sum(apply(xx<2, MARGIN=1, FUN=sum)) #two trials with no choice (and hence no latency)
which(latency_fc_data$latency>120) # one trial where the dog needed more than 2 minutes to steal the treat, hence we consider it as "no choice" and exclude it from the analyses
latency_fc_data2<-latency_fc_data[-220,]
max(latency_fc_data2$latency)
min(latency_fc_data2$latency)

#divide latencies by 120 (2 min=max trial duration)
latency_fc_data2$prop.latency<-latency_fc_data2$latency/120
max(latency_fc_data2$prop.latency)
min(latency_fc_data2$prop.latency) #Distribution does not include 0s and 1s
hist(latency_fc_data2$prop.latency)
```


```{r}
#scale and center covariate
latency_fc_data2$z.trial<-as.vector(scale(latency_fc_data2$trial_w_con, center = TRUE, scale=TRUE))
#Center factor
latency_fc_data2$condition.c<-as.vector(scale(as.numeric(latency_fc_data2$condition==levels(as.factor(latency_fc_data2$condition))[2]), center=TRUE, scale= FALSE))
```
#### fit beta model
```{r}
library(glmmTMB)
mm1.latency.int<-glmmTMB(prop.latency ~ choice*condition + z.trial + first_condition + 
          (1 + condition.c | Subject),
            family=beta_family, data=latency_fc_data2,  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))

#Warning: Model convergence problem; non-positive-definite Hessian matrix. 
#Hence, removed correlation between random slope and intercept

mm1.latency.int<-glmmTMB(prop.latency ~ choice*condition + z.trial + first_condition + 
          (1 | Subject)+
          (0 + condition.c |Subject),
            family=beta_family, data=latency_fc_data2,  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))


#check significance of interaction
drop1_mm1.latency.int<-drop1(mm1.latency.int, test="Chisq")%>%filter(Df!="")%>%add_row(Df = rep(NA,1),  .before = 1)
drop1_mm1.latency.int
#interaction not significant (Chisq.1 = 0.46, p = 0.499). Hence, dropped from model:

mm1.latency<-glmmTMB(prop.latency ~ choice + condition + z.trial + first_condition + 
          (1 | Subject)+
          (0 + condition.c |Subject),
            family=beta_family, data=latency_fc_data2,  control=glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000)))

#check for absence of overdispersion 
overdisp.test(mm1.latency) #disp. parameter= 1.07, ok.
```
####Model output
```{r}
#coefficients
summary(mm1.latency)$coeff$cond[,1:2]
#tests
drop1_mm1.latency<-drop1(mm1.latency, test="Chisq")%>%filter(Df!="")%>%add_row(Df = rep(NA,1),  .before = 1)
drop1_mm1.latency

mm1_latency_aic<-drop1(mm1.latency, test="Chisq")[1, "AIC"]
mm1_latency_aic_drop1<-drop1(mm1.latency, test="Chisq")[, "AIC"][-1]
mm1_latency_aic_drop1_delta<- c(NA, round(mm1_latency_aic - mm1_latency_aic_drop1, 2))
mm1_latency_aic_drop1_delta_df <- data.frame(delta_aic = mm1_latency_aic_drop1_delta)
```
####Model stability

```{r}
mm1.latency.stab=glmmTMB.stab(model.res=mm1.latency, para=F, data=latency_fc_data2)

mm1.latency.stab$summary[1:5, -1]

m.stab.plot(round(mm1.latency.stab$summary[1:5, -1], 2))
#is.re=grepl(x=rownames(mm1.beta.stab$summary), pattern="@")
#m.stab.plot(mm1.beta.stab$summary[!is.re, -1])
```
#####collinearity
```{r}
library(car)
xx=lm(prop.latency~condition+choice+z.trial+first_condition, data=latency_fc_data2_model)
vif(xx)
```
#### confidence intervals

```{r}
boot.res.latency=boot.glmmtmb(mm1.latency, data=latency_fc_data2, nboots=1000, para=T, n.cores = "all-1", resol=100, level=0.95)

boot.res.latency$ci.estimates$fe[1:5,]
```
#### output table - latency beta model
```{r}
mm1_beta_latency_output_table <-
  bind_cols(as.data.frame(summary(mm1.latency)$coeff$cond[,1:2]),
            drop1_mm1.latency,
            mm1.latency.stab$summary[1:5, -1],
            boot.res.latency$ci.estimates$fe[1:5,],
            mm1_latency_aic_drop1_delta_df) %>%
  dplyr::select(
    Estimate,
    SE = `Std. Error`,
    Chi2 = LRT,
    Df,
    p = `Pr(>Chi)`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    min,
    max,
    delta_aic
  ) %>% 
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Df), ~ format(round(.x, 2), nsmall = 2))) %>%
  mutate(across(.cols = c(LowerCI:max), ~ format(round(.x, 2), nsmall = 2))) %>% 
  dplyr::mutate(p = replace(p, p == "0", "<.001"))


write.csv(mm1_beta_latency_output_table, file = "saves/mm1_beta_latency_output_table.csv")
```
## Peeking

```{r}
peeking_data <- loopy.data %>%
  filter(Behaviour == "peeking")

levels(as.factor(peeking_data$Value)) #dogs never peeked behind the opaque tunnel
```






## second choice latency

#### Model preparation
```{r}
latency_sc_data<- latency_sc_data%>%
  dplyr::select(Subject, latency, condition, trial_w_con, first_condition)%>%
  full_join(first_choice_data%>%dplyr::select(Subject, opaque_choice, condition, trial_w_con, first_condition))%>%
  mutate(choice = fct_recode(as.factor(opaque_choice), "opaque" ="1", "clear" ="0"))



latency_sc_data$z.trial<-as.vector(scale(latency_sc_data$trial_w_con, center = TRUE, scale=TRUE))
latency_sc_data$condition <- as.factor(latency_sc_data$condition)
levels(latency_sc_data$condition)
latency_sc_data$condition.c=as.vector(scale(as.numeric(latency_sc_data$condition==levels(as.factor(latency_sc_data$condition))[2]), center=TRUE, scale= FALSE))

latency_sc_data$choice.c=as.vector(scale(as.numeric(latency_sc_data$choice==levels(as.factor(latency_sc_data$choice))[2]), center=TRUE, scale= FALSE))




latency_sc_data$log_latency<-log(latency_sc_data$latency)
hist(latency_sc_data$latency)
max(latency_sc_data$latency, na.rm=TRUE)
min(latency_sc_data$latency, na.rm=TRUE)
is.na(latency_sc_data$latency)
hist(latency_sc_data$log_latency)

table(latency_sc_data$Subject, latency_sc_data$choice)

latency_sc_data<-latency_sc_data%>%filter(!is.na(latency))
```




#### fit model
```{r}
mm1.sc.latency.int <- lme4::lmer(
  log_latency ~ choice*condition + z.trial + first_condition +
    (1 + condition.c | Subject),
  data = latency_sc_data,
  REML = FALSE
)
summary(mm1.sc.latency.int)
drop1(mm1.sc.latency.int, test="Chisq") # interaction between choice and condition  significant 
```



```{r}
library(performance)
#check_model(mm1.sc.latency.reml)
diagnostics.plot(mm1.sc.latency.int, size.fac=2)

ranef.diagn.plot(mm1.sc.latency.int)
```
```{r}
ggplot(latency_sc_data, aes(x= choice, y = log_latency, color=condition))+

  geom_boxplot()+
    geom_point(position=position_dodge(width = 0.2), alpha=.3)+
  facet_wrap(~trial_w_con)
```

```{r}
latency_sc_data_agg<-latency_sc_data%>%
  group_by(condition, choice)%>%
  summarise(mean_latency = mean(latency, na.rm=TRUE), se=sd(latency, na.rm=TRUE)/sqrt(length(latency)))

sc.plot<-ggplot(data = latency_sc_data_agg, aes(x = as.factor(choice), y= mean_latency, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_latency - se, ymax = mean_latency + se  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  xlab(NULL) + 
  ylab ("Mean second choice latency")+
  scale_fill_manual(values = c("darkorange", "darkgreen"))

ggsave(sc.plot, file="graphics/second_choice_latency_plot.png", width = 6, height= 6, scale=0.8)

latency_sc_data_agg2<-latency_sc_data%>%
  group_by(condition, choice, trial_w_con)%>%
  summarise(mean_latency = mean(latency, na.rm=TRUE), se=sd(latency, na.rm=TRUE)/sqrt(length(latency)))

ggplot(data = latency_sc_data_agg2, aes(x = as.factor(choice), y= mean_latency, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_latency - se, ymax = mean_latency + se  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  xlab(NULL) + 
  ylab ("Mean second choice latency")+
  scale_fill_manual(values = c("darkorange", "darkgreen"))+
  facet_wrap(~trial_w_con)

```


```{r}
drop1_mm1.sc.latency <- drop1(mm1.sc.latency.int, test="Chisq")%>%
  filter(!is.na(npar))%>%
  add_row(npar = rep(NA,1),  .before = 1)
drop1_mm1.sc.latency

```
Individual fixed effects
```{r}
library(lmerTest)

mm1.sc.latency.reml<-lmerTest::lmer(
  log_latency ~ choice * condition + z.trial + first_condition +
    (1 + condition.c | Subject),
  data = latency_sc_data,
  REML=TRUE)

summary(mm1.sc.latency.reml)
```
+ Relevel the reference category of choice
```{r}
library(lmerTest)
latency_sc_data$choice_rel <- relevel( as.factor(latency_sc_data$choice), ref = "opaque") 
latency_sc_data$condition_rel <- relevel( as.factor(latency_sc_data$condition), ref = "test") 
levels(latency_sc_data$choice_rel)
mm1.sc.latency.reml_rel<-lmerTest::lmer(
  log_latency ~ choice_rel * condition + z.trial + first_condition +
    (1 + condition.c | Subject),
  data = latency_sc_data,
  REML=TRUE)

summary(mm1.sc.latency.reml_rel)
```


Colinearity checks
```{r}
library(car)
xx=lm(log_latency ~ choice + condition + z.trial + first_condition,
  data = latency_sc_data)
vif(xx)
```
relative model complexity
```{r}
length(residuals(mm1.sc.latency))/
(length(fixef(mm1.sc.latency))+
nrow(as.data.frame(summary(mm1.sc.latency)$varcor)))
```
model stability
To evaluate model stability we removed each level of the random intercept subject ID one at a time and compared the resulting estimate. This revealed the model to be stable with respect to the fixed effects.
```{r}

mm1.sc.latency.stab=glmm.model.stab(model.res=mm1.sc.latency.int, contr=NULL, para=F, data=NULL)

mm1.sc.latency.stab$summary

m.stab.plot(round(mm1.sc.latency.stab$summary[, -1], 3))
```
--> model stable with regard to the fixed effects

```{r}
boot.mm1.sc.latency=boot.lmer(mm1.sc.latency.int,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)

mm1.sc.latency_boot_ci<-boot.mm1.sc.latency$ci.estimates
```
#### effect size
```{r}
library(MuMIn)
r.squaredGLMM(mm1.sc.latency.int)
```


#### output table
```{r}


model_table_mm1.sc.latency <- bind_cols(as.data.frame(summary(mm1.sc.latency.reml)$coefficients),

                         mm1.sc.latency_boot_ci) %>% #mm2_boot_ci<-boot.full$ci.estimates
  dplyr::select(Estimate, SE = `Std. Error`, LowerCI = X2.5., UpperCI = X97.5., t_value = `t value`, p = `Pr(>|t|)`) %>%
  mutate(across(.cols = c(p), ~ round(.x, 3))) %>% 
  mutate(across(.cols = c(Estimate:t_value), ~ round(.x, 2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
mutate(p=replace(p, p==0, "<0.001"))

write.csv(model_table_mm1.sc.latency , file = "saves/sc_latency_mm1_output_table.csv")
```

##LL: used from here until line 817
## subgroups: terrier, etc
### terrier
#### choice(terrier)
aggregate data

```{r}
terrier_first_choice_data.agg <- first_choice_data %>%
  mutate(terrier = fct_recode(as.factor(terrier), "Terrier" = "yes", "No terrier" = "no"))%>%
  group_by(Subject, condition, terrier, trial_w_con) %>%
  summarise(mean_opaque_ind = mean(opaque_choice)) 

terrier_first_choice_data.agg %>%   group_by(trial_w_con, condition, terrier) %>% summarise(length(mean_opaque_ind))

terrier_first_choice_data.agg2 <- terrier_first_choice_data.agg %>%
  group_by(condition, terrier, trial_w_con) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))


terrier_first_choice_data.agg3 <- terrier_first_choice_data.agg %>%
  group_by(condition, terrier) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))
```



```{r}
ggplot(data = terrier_first_choice_data.agg2, aes(x = as.factor(trial_w_con), y= mean_opaque, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  ylim(0, 1) + 
  xlab("Trial within condition") + 
  ylab ("Mean proportion opaque choice")+
  facet_wrap(~terrier)+ 
  scale_fill_manual(values = c("darkorange", "darkgreen"))

```

```{r}
ggplot(data = terrier_first_choice_data.agg3, aes(x =terrier, y= mean_opaque, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  ylim(0, 1) + 
  xlab(NULL) + 
  ylab ("Mean proportion opaque choice")+ 
  scale_fill_manual(values = c("darkorange", "darkgreen"))+
  geom_hline(yintercept = 0.5, lty = 2, color = "grey")
```

fit model


```{r}
mm1_choice_terrier=glmer(opaque_choice ~ terrier*condition+z.trial+first_condition+
                   (1+condition.c|Subject),
             data=first_choice_data, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mm1_choice_terrier)
drop1(mm1_choice_terrier, test="Chisq") #interaction not significant
```
```{r}
mm2_choice_terrier=glmer(opaque_choice ~ terrier+condition+z.trial+first_condition+
                   (1+condition.c|Subject),
             data=first_choice_data, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mm2_choice_terrier)
drop1(mm2_choice_terrier, test="Chisq")
```

```{r}

mm2_choice_terrier_drop1 <- drop1(mm2_choice_terrier, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1)
```

Check for colinearity

```{r}
library(car)
xx=lm(opaque_choice ~ terrier+condition+z.trial+first_condition, data=first_choice_data)
vif(xx)
```

```{r}
mm2_choice_terrier.ci=boot.glmm.pred(model.res=mm2_choice_terrier, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)

```
#### stability (terrier)
```{r}
m.stab.fc.terrier <- glmm.model.stab(model.res = mm2_choice_terrier)
m.stab.fc.terrier$detailed$warnings
as.data.frame(round(m.stab.fc.terrier$summary[1:5, -1], 2))

m.stab.plot(round(m.stab.fc.terrier$summary[, -1], 2))
```

#### output table

```{r}
mm2_terrier_choice_output_table <-
  bind_cols(as.data.frame(summary(mm2_choice_terrier)$coeff),
            mm2_choice_terrier_drop1,
            mm2_choice_terrier.ci$ci.estimates,
            m.stab.fc.terrier$summary[1:5, -1]) %>%
   dplyr::select(
    Estimate,
    SE = `Std. Error`,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    min,
    max
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:df), ~ format(round(.x, 2), nsmall = 2))) %>%
  mutate(across(.cols = c(LowerCI:max), ~ format(round(.x, 2), nsmall = 2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))


write.csv(mm2_terrier_choice_output_table, file = "saves/mm2_choice_terrier_output_table.csv")
```



### latency data (terrier)

```{r}

latency_fc_data_terrier <- loopy.data%>%
  mutate(Value = ifelse(Subject == "Sophi" & Start_Frame == 143869, "latency_to_2nd_choice", Value))%>%
  filter(Behaviour == "latency_first_choice", Value == "latency") %>%
  rename(latency = duration)%>%
  mutate(terrier = fct_recode(as.factor(terrier), "Terrier" = "yes", "No terrier" = "no"))%>%
  select(-opaque_choice)%>%
  full_join(first_choice_data%>%select(Subject, opaque_choice, condition, trial_w_con))%>%
  mutate(choice = fct_recode(as.factor(opaque_choice), "opaque" ="1", "clear" ="0"))
  



```

aggregate latency data

```{r}
latency_fc_data_terrier_agg <- latency_fc_data_terrier %>%
  filter(Value =="latency")%>%
  group_by(Subject, condition, choice, terrier) %>%
  summarise(mean_latency = mean(latency))

latency_data_agg2 <- latency_fc_data_terrier_agg %>%
  group_by(condition, choice, terrier) %>%
  summarise(mean_latency2 = mean(mean_latency), se_latency = sd(mean_latency)/sqrt(length(mean_latency)))
```

```{r}
ggplot(latency_fc_data_terrier_agg, aes(x = condition, y = mean_latency)) + 
  geom_boxplot(outlier.colour = "white")+
  geom_jitter()+
  facet_grid(choice~terrier)+ 
  theme_bw()
```

```{r}
ggplot(latency_data_agg2, aes(x = condition, y = mean_latency2, fill=choice)) +
  geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_latency2 - se_latency, ymax = mean_latency2 + se_latency  ),  position=position_dodge(.9), width=0.4)+
  facet_wrap(~terrier)+ 
  theme_bw()

```
#### Model preparation
```{r}
latency_fc_data_terrier2 <- latency_fc_data_terrier2 %>%
  filter(terrier =="Terrier")
latency_fc_data_terrier2$z.trial<-as.vector(scale(latency_fc_data_terrier2$trial_w_con, center = TRUE, scale=TRUE))
latency_fc_data_terrier2$condition <- as.factor(latency_fc_data_terrier2$condition)
levels(latency_fc_data_terrier2$condition)
latency_fc_data_terrier2$condition.c=as.vector(scale(as.numeric(latency_fc_data_terrier2$condition==levels(as.factor(latency_fc_data_terrier2$condition))[2]), center=TRUE, scale= FALSE))

latency_fc_data_terrier2$choice.c=as.vector(scale(as.numeric(latency_fc_data_terrier2$choice==levels(as.factor(latency_fc_data_terrier2$choice))[2]), center=TRUE, scale= FALSE))

latency_fc_data_terrier2$log_latency<-log(latency_fc_data_terrier2$latency)
hist(latency_fc_data_terrier2$latency)
max(latency_fc_data_terrier2$latency)

hist(latency_fc_data_terrier2$log_latency)

table(latency_fc_data_terrier2$Subject, latency_fc_data_terrier2$choice)
```




#### fit model
```{r}
mm1.latency.terrier.int <- lme4::lmer(
  log_latency ~ choice*condition + z.trial + first_condition +
    (1 + condition.c | Subject),
  data = latency_fc_data_terrier2,
  REML = FALSE
)
summary(mm1.latency.terrier.int)
drop1(mm1.latency.terrier.int, test="Chisq") # interaction between choice and condition not significant --> interaction is dropped to evaluate the main effects
```
```{r}
mm1.latency.terrier <- lme4::lmer(
  log_latency ~ choice+condition + z.trial + first_condition +
    (1 + condition.c | Subject),
  data = latency_fc_data_terrier2,
  REML = FALSE
)
summary(mm1.latency.terrier)
drop1(mm1.latency.terrier, test="Chisq")
```


```{r}
diagnostics.plot(mm1.latency.terrier, size.fac=2)

ranef.diagn.plot(mm1.latency.terrier)
```

```{r}
drop1_mm1.latency.terrier <- drop1(mm1.latency.terrier, test="Chisq")%>%
  filter(!is.na(npar))%>%
  add_row(npar = rep(NA,1),  .before = 1)
drop1_mm1.latency.terrier

```
Individual fixed effects
```{r}
library(lmerTest)

mm1.latency.terrier.reml<-lmerTest::lmer(
  log_latency ~ choice + condition + z.trial + first_condition +
    (1 + condition.c | Subject),
  data = latency_fc_data_terrier2,
  REML=TRUE)

summary(mm1.latency.terrier.reml)
```



Colinearity checks
```{r}
library(car)
xx=lm(log_latency ~ choice + condition + z.trial + first_condition,
  data = latency_fc_data_terrier2)
vif(xx)
```
relative model complexity
```{r}
length(residuals(mm1.latency.terrier))/
(length(fixef(mm1.latency.terrier))+
nrow(as.data.frame(summary(mm1.latency.terrier)$varcor)))
```
model stability
To evaluate model stability we removed each level of the random intercept subject ID one at a time and compared the resulting estimate. This revealed the model to be stable with respect to the fixed effects.
```{r}

mm1.latency.terrier.stab=glmm.model.stab(model.res=mm1.latency.terrier, contr=NULL, para=F, data=NULL)

mm1.latency.terrier.stab$summary

m.stab.plot(round(mm1.latency.terrier.stab$summary[, -1], 3))
```
--> model stable with regard to the fixed effects

```{r}
boot.mm1.latency.terrier=boot.lmer(mm1.latency.terrier,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)

mm1.latency.terrier_boot_ci<-boot.mm1.latency.terrier$ci.estimates
```
#### effect size
```{r}
library(MuMIn)
r.squaredGLMM(mm1.latency.terrier)
```


#### output table
```{r}


model_table_mm1.latency.terrier <- bind_cols(as.data.frame(summary(mm1.latency.terrier.reml)$coefficients),
                         #drop1_mm1.yarbus.signaling,
                         mm1.latency.terrier_boot_ci) %>% #mm2_boot_ci<-boot.full$ci.estimates
  dplyr::select(Estimate, SE = `Std. Error`, LowerCI = X2.5., UpperCI = X97.5., t_value = `t value`, p = `Pr(>|t|)`) %>%
  mutate(across(.cols = c(p), ~ round(.x, 3))) %>% 
  mutate(across(.cols = c(Estimate:t_value), ~ round(.x, 2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
mutate(p=replace(p, p==0, "<0.001"))

write.csv(model_table_mm1.latency.terrier , file = "saves/latency_mm1_terrier_output_table.csv")
```


#First trial only  - analysis of first choice
```{r}
#the dog that needed more than 120 s to make his first choice in his first trial (but started moving within 120 s) is included in this analysis.
first_choice_data_with60subjcs <- loopy.data %>%
  filter(Behaviour == "first_choice")

first_trial_data<-first_choice_data_with60subjcs %>% 
  filter(trial==1)
length(levels(as.factor(first_trial_data$Subject)))
```
```{r}
#prepare data for model
#center factor
first_trial_data$condition <- as.factor(first_trial_data$condition)
levels(first_trial_data$condition)
first_trial_data$condition.c=as.vector(scale(as.numeric(first_trial_data$condition==levels(as.factor(first_trial_data$condition))[2]), center=TRUE, scale= FALSE))

```

```{r}
#fit model
mm1_choice_first_trial=glm(opaque_choice ~ condition,
             family=binomial,
             data=first_trial_data)
summary(mm1_choice_first_trial)
drop1(mm1_choice_first_trial, test="Chisq")
```

```{r}
#model stability

my.dfbeta<-function(m){
xx=cbind(coef(m), coef(m)+
t(apply(X=dfbeta(m), MARGIN=2, FUN=range)))
colnames(xx)=c("orig", "min", "max")
return(xx)
}
mm1_first_trial_stab<-my.dfbeta(mm1_choice_first_trial)
mm1_first_trial_stab[,-1]
```
```{r}
#confidence intervals
cbind(orig=coef(mm1_choice_first_trial), confint(object=mm1_choice_first_trial))
cis.first.trial<-confint(object=mm1_choice_first_trial)
```


```{r}
#results
round(summary(mm1_choice_first_trial)$coefficients,3)
```

```{r}
#output table - first trial
mm1_choice_first_trial_output_table <-
  bind_cols(as.data.frame(summary(mm1_choice_first_trial)$coefficients),
            cis.first.trial,
            mm1_first_trial_stab[,-1]) %>%
  dplyr::select(
    Estimate,
    SE = `Std. Error`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`,
    LowerCI = `2.5 %`,
    UpperCI = `97.5 %`,
    min,
    max
  ) %>% #
  mutate(across(.cols = c(p_wald), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:z_wald), ~ format(round(.x, 2), nsmall = 2))) %>%
  mutate(across(.cols = c(LowerCI:max), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p_wald = replace(p_wald, p_wald == 0, "<0.001"))

write.csv(mm1_choice_first_trial_output_table, file = "saves/mm1_choice_first_trial_output_table.csv")
```

### independent breeds

#### According to Heberlein
aggregate data

```{r}
ind_first_choice_data.agg <- first_choice_data %>%
  mutate(independent_heberlein = fct_recode(as.factor(independent_heberlein), "Independent" = "yes", "Non-independent" = "no"))%>%
  group_by(Subject, condition, independent_heberlein, trial_w_con) %>%
  summarise(mean_opaque_ind = mean(opaque_choice)) 

ind_first_choice_data.agg %>%   group_by(trial_w_con, condition, independent_heberlein) %>% summarise(length(mean_opaque_ind))
table(ind_first_choice_data.agg$condition, ind_first_choice_data.agg$independent_heberlein )


ind_first_choice_data.agg2 <- ind_first_choice_data.agg %>%
  group_by(condition, independent_heberlein, trial_w_con) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))


ind_first_choice_data.agg3 <- ind_first_choice_data.agg %>%
  group_by(condition, independent_heberlein) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))


breed_first_choice_data.agg <- first_choice_data %>%
  mutate(independent_heberlein = fct_recode(as.factor(independent_heberlein), "Independent" = "yes", "Non-independent" = "no"))%>%
  group_by(Subject, independent_heberlein, breed) %>%
  summarise(n = length(trial==1))

table(breed_first_choice_data.agg$breed, breed_first_choice_data.agg$independent_heberlein)
```

```{r}
ggplot(data = ind_first_choice_data.agg2, aes(x = as.factor(trial_w_con), y= mean_opaque, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  ylim(0, 1) + 
  xlab("Trial within condition") + 
  ylab ("Mean proportion opaque choice")+
  facet_wrap(~independent_heberlein)+ 
  scale_fill_manual(values = c("darkorange", "darkgreen"))

```

```{r}
ggplot(data = ind_first_choice_data.agg3, aes(x =independent_heberlein, y= mean_opaque, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  ylim(0, 1) + 
  xlab(NULL) + 
  ylab ("Mean proportion opaque choice")+ 
  scale_fill_manual(values = c("darkorange", "darkgreen"))+
  geom_hline(yintercept = 0.5, lty = 2, color = "grey")
```

#### According to FCI and Heberlein
aggregate data

```{r}
indfci_first_choice_data.agg <- first_choice_data %>%
  filter(independent_breed_fci!="")%>%
  mutate(independent_breed_fci = fct_recode(as.factor(independent_breed_fci), "Independent" = "yes", "Family-style" = "no"))%>%
  group_by(Subject, condition, independent_breed_fci, trial_w_con) %>%
  summarise(mean_opaque_ind = mean(opaque_choice)) 

indfci_first_choice_data.agg %>%   group_by(trial_w_con, condition, independent_breed_fci) %>% summarise(length(mean_opaque_ind))

ind_first_choice_data.agg2 <- indfci_first_choice_data.agg %>%
  group_by(condition, independent_breed_fci, trial_w_con) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))


ind_first_choice_data.agg3 <- indfci_first_choice_data.agg %>%
  group_by(condition, independent_breed_fci) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))


breed_first_choice_data.agg <- first_choice_data %>%
    filter(independent_breed_fci!="")%>%
  mutate(independent_fci = fct_recode(as.factor(independent_breed_fci), "Independent" = "yes", "Family-style" = "no"))%>%
  group_by(Subject, independent_fci, breed) %>%
  summarise(n = length(trial==1))

table(breed_first_choice_data.agg$breed, breed_first_choice_data.agg$independent_fci)
```

```{r}
ggplot(data = ind_first_choice_data.agg2, aes(x = as.factor(trial_w_con), y= mean_opaque, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque  ),  position=position_dodge(.9), width=0.4)+
  theme_bw() +
  ylim(0, 1) + 
  xlab("Trial within condition") + 
  ylab ("Mean proportion opaque choice")+
  facet_wrap(~independent_breed_fci)+ 
  scale_fill_manual(values = c("darkorange", "darkgreen"))

```

#### According to FCI groups
aggregate data

```{r}
fci_first_choice_data.agg <- first_choice_data %>%
    mutate(
    fci_group = fct_recode(
      as.factor(fci_group),
      "Pinscher and Schnauzer" = "2",
      "Terriers" = "3",
      "Dachshunds" = "4",
      "Spitz and primitive types" = "5",
      "Retrievers" = "8",
      "Companion and Toy Dogs" = "9",
      "Sighthounds" = "10"
          )
  )%>%
  group_by(Subject, condition, fci_group, trial_w_con) %>%
  summarise(mean_opaque_ind = mean(opaque_choice)) 

fci_first_choice_data.agg %>%   group_by(trial_w_con, condition, fci_group) %>% summarise(length(mean_opaque_ind))

fci_first_choice_data.agg2 <- fci_first_choice_data.agg %>%
  group_by(condition, fci_group, trial_w_con) %>%
  summarise(mean_opaque = mean(mean_opaque_ind), median_opaque = median(mean_opaque_ind), sd_opaque = sd(mean_opaque_ind), se_opaque = sd(mean_opaque_ind)/sqrt(length(mean_opaque_ind)))



fci_first_choice_data.agg3 <- fci_first_choice_data.agg %>%
  group_by(condition, fci_group) %>%
  summarise(
    mean_opaque = mean(mean_opaque_ind, na.rm=TRUE),
    median_opaque = median(mean_opaque_ind),
    sd_opaque = sd(mean_opaque_ind, na.rm=TRUE),
    se_opaque = sd(mean_opaque_ind, na.rm=TRUE) / sqrt(length(mean_opaque_ind)),
    sample_size=length(mean_opaque_ind)
  )

fci_first_choice_data.agg4 <- fci_first_choice_data.agg %>%
  group_by(fci_group) %>%
  summarise(
        mean_opaque = mean(mean_opaque_ind, na.rm=TRUE),
    median_opaque = median(mean_opaque_ind),
    sd_opaque = sd(mean_opaque_ind, na.rm=TRUE),
    se_opaque = sd(mean_opaque_ind, na.rm=TRUE) / sqrt(length(mean_opaque_ind)),
    sample_size=sum(trial_w_con==1 & condition=="test")
  )

fci_breed_first_choice_data.agg <- first_choice_data %>%
  mutate(
    fci_group = fct_recode(
      as.factor(fci_group),
      "Pinscher and Schnauzer" = "2",
      "Terriers" = "3",
      "Dachshunds" = "4",
      "Spitz and primitive types" = "5",
      "Retrievers" = "8",
      "Companion and Toy Dogs" = "9",
      "Sighthounds" = "10"
          )
  ) %>%
  group_by(Subject, fci_group, breed) %>%
  summarise(n = length(trial == 1))

table(fci_breed_first_choice_data.agg$breed, fci_breed_first_choice_data.agg$fci_group)
```

```{r}
ggplot(data = fci_first_choice_data.agg2, aes(x = as.factor(trial_w_con), y= mean_opaque, fill=condition))+
    geom_bar(stat = "identity", position=position_dodge(.9))+
  geom_errorbar(aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque  ),  position=position_dodge(.9), width=0.4)+
  geom_hline(yintercept=0.5, lty=2)+
  theme_bw() +
  ylim(0, 1) + 
  xlab("Trial within condition") + 
  ylab ("Mean proportion opaque choice")+
  facet_wrap(~fci_group)+ 
  scale_fill_manual(values = c("darkorange", "darkgreen"))

```
```{r}

fci_plot <-
  ggplot(data = fci_first_choice_data.agg3 %>% filter(!fci_group %in% c("Dachshunds", "Retrievers")),
         aes(x = fci_group, y = mean_opaque, fill = condition)) +
  geom_bar(stat = "identity", position = position_dodge(.9)) +
  geom_errorbar(
    aes(ymin = mean_opaque - se_opaque, ymax = mean_opaque + se_opaque),
    position = position_dodge(.9),
    width = 0.4
  ) +
  theme_bw() +
  coord_flip() +
  ylim(0, 1) +
  xlab(NULL) +
  ylab ("Mean proportion opaque choice") +
  scale_fill_manual(values = c("darkorange", "darkgreen")) +
  geom_hline(yintercept = 0.5,
             lty = 2,
             color = "grey") +
  geom_text(
    data = fci_first_choice_data.agg4 %>% filter(!fci_group %in% c("Dachshunds", "Retrievers")),
    aes(label = paste0("n = ", sample_size), fill = NULL),
    position = position_dodge(width = 0),
    hjust = -2,
    size = 4
  )
fci_plot
```
```{r}
ggsave(fci_plot, file = "graphics/fci_plot.png", width = 10, height = 6, scale = 0.7)
```

### terrier and primitive types
aggregate data

```{r}
fci3_5_first_choice_data <- first_choice_data %>%
  filter(fci_group %in% c("3", "5"))

```


```{r}
fci3_5_first_choice_data$z.trial<-as.vector(scale(fci3_5_first_choice_data$trial_w_con, center = TRUE, scale=TRUE))

fci3_5_first_choice_data$condition <- as.factor(fci3_5_first_choice_data$condition)
levels(fci3_5_first_choice_data$condition)
fci3_5_first_choice_data$condition.c=as.vector(scale(as.numeric(fci3_5_first_choice_data$condition==levels(as.factor(fci3_5_first_choice_data$condition))[2]), center=TRUE, scale= FALSE))
```

```{r}

fci35_first_choice_data.agg <- fci3_5_first_choice_data %>%
  group_by(condition) %>%
  summarise(
    mean_opaque = mean(opaque_choice, na.rm=TRUE),
    median_opaque = median(opaque_choice),
    sd_opaque = sd(opaque_choice, na.rm=TRUE),
    se_opaque = sd(opaque_choice, na.rm=TRUE) / sqrt(length(opaque_choice)),
    sample_size=length(opaque_choice)
  )

ggplot(fci35_first_choice_data.agg, aes(x = condition, y = mean_opaque))+
  geom_bar(stat="identity")
```


fit model

```{r}
mm1_choice_fcigr=glmer(opaque_choice ~ condition+z.trial+first_condition+
                   (1+condition.c|Subject),
             data=fci3_5_first_choice_data, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mm1_choice_fcigr)
drop1(mm1_choice_fcigr, test="Chisq") 
```

### pretest carry over effect

fit model

```{r}
first_choice_data$firstchoice_pretest2<-trimws(first_choice_data$firstchoice_pretest2)
first_choice_data<-first_choice_data%>%
  mutate(firstchoice_pretest2 = fct_recode(as.factor(firstchoice_pretest2), "transparent" = "transparant"))
levels(as.factor(first_choice_data$firstchoice_pretest2))
mm1_choice_pretest=glmer(opaque_choice ~ firstchoice_pretest2+condition+z.trial+first_condition+
                   (1+condition.c|Subject),
             data=first_choice_data, family=binomial, 
            control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mm1_choice_pretest)
drop1(mm1_choice_pretest, test="Chisq")
```

```{r}

mm1_choice_pretest_drop1 <- drop1(mm1_choice_pretest, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1)
```

Check for colinearity

```{r}
library(car)
xx=lm(opaque_choice ~ firstchoice_pretest2+condition+z.trial+first_condition,
                    data=first_choice_data)
vif(xx)
```

```{r}
mm1_choice_pretest.ci=boot.glmm.pred(model.res=mm1_choice_pretest, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)

```

#### output table

```{r}
mm1_choice_pretest_output_table <-
  bind_cols(as.data.frame(summary(mm1_choice_pretest)$coeff),
            mm1_choice_pretest_drop1,
            mm1_choice_pretest.ci$ci.estimates) %>%
  dplyr::select(
    Estimate,
    SE = `Std. Error`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_choice_pretest_output_table, file = "saves/mm1_choice_pretest_output_table.csv")
```