---
title: "Dogs stealing behind tunnels: power simulation"
author: "Christoph Völter"
date: "15/09/2022"
output: 
  html_document:
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(cowplot)
library("gghalves")
library(ggthemes)

load("stealing_behind_tunnels_80subj.RData")
```

Notes:
DV: choice of opaque side
Design: 2 conditions (test / control), within subject design
1 session, 1 trials per condition (increase if needed)
Prediction: opaque side in 65% of test trials, and 50% in control trials
N = 80

## Generate data
```{r echo=FALSE, include=FALSE}
set.seed(1)
n.subject <- 80 # number subjects
n.per.subject <- 2 # observations per subject
n.per.condition <- 1 # observations per subject and condition
n.sessions <- 1
#age_range <- c(12:130) # age range between 1 and 13 years
test.per<-c(0.65, 0.7) # test condition
control.per<-c(0.5) # control condition

subj.id <- as.factor(paste("subj", str_pad(1:n.subject, 2, pad = "0"), sep = "."))
start.data <- data.frame(expand.grid(subj.id = subj.id, condition = c("test", "control")))
start.data <- arrange(start.data, subj.id)
start.data$trial <- c(rep(c(1,2),n.subject/2), rep(c(2,1),n.subject/2))


start.data$z.trial <- as.vector(scale(as.numeric(start.data$trial)))
#start.data$z.session <- as.vector(scale(as.numeric(start.data$session)))

# dummy code factors
start.data$condition <- as.factor(start.data$condition)
start.data$condition <- relevel(start.data$condition, ref = "control")
levels(start.data$condition)
start.data$condition.dummy <- as.numeric(start.data$condition == levels(start.data$condition)[2])

# center condition for random slopes:
start.data$condition.c <- as.numeric(start.data$condition) - mean(as.numeric(start.data$condition))
```

data checks
```{r echo=FALSE, include=FALSE}
table(start.data$subj.id, start.data$condition)
table(start.data$trial, start.data$condition)
```

## Simulation

```{r eval=FALSE, include=FALSE}
n.simus <- 1000
r.effects <- c(0.62, 1.24) # random effects to be simulated
# with the coefficient of condition being 0.62 (qlogis(0.65)) we assume a moderately large random intercept of 0.62.

#r.slope.int <- c(1.1)
# with the estimate being -1.098612 (qlogis(0.25)-qlogis(0.5)) we assume a moderately large random slope of 1.1.

#r.slope.trial <- 0.2

# create object to store the simulation parameters and results:
all.res <- data.frame(expand.grid(
  n.per.subject = n.per.subject, r.effect = r.effects,
 # r.slope.int = r.slope.int, r.slope.trial = r.slope.trial,
  #r.slope.session = r.slope.session,
  test.per = test.per,
  control.per = control.per,
  simu = 1:n.simus
))
all.res$icpt <- NA
all.res$conditiontest <- NA
all.res$re.sd <- NA
all.res$warns.full <- NA
all.res$warns.null <- NA
all.res$lrt.p.con <- NA
#all.res$lrt.p.age <- NA
all.res$full.null.p <- NA

all.ests <- matrix(NA, nrow = n.simus, ncol = 1)
colnames(all.ests) <- c("lrt.p.con")

# create data frame with design:
## done above

# load packages needed:
library(lme4)
# Loading required package: Matrix
library(kyotil) # we want to store info about convergence issues

# define control structure to make convergence more likely:
contr <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))

xdata <- start.data
m.mat <- model.matrix(object = ~condition + z.trial, data = xdata) # create model martix

# run simulation
for (i in 1:nrow(all.res)) {

  set.seed(i) # allows to later replicate individual simulations
  # xdata <- start.data[rep(x = 1:nrow(start.data), each = all.res[i, "n.per.subject"]), ]  # replicate rows in start.data according to sample size
  # add age  (if it should be generated in each loop)
  #age <- sample(x = age_range, size = length(unique(xdata$subj.id)), replace = T)
  #xdata$age <- as.numeric(age[as.numeric(xdata$subj.id)])
  #xdata$z.age <- scale(xdata$age)
  #m.mat <- model.matrix(object = ~condition + z.age + sex + z.trial + z.session , data = xdata) # create model martix

  coefs <- c(
    "(Intercept)" = qlogis(all.res[i, "control.per"]),
    "conditiontest" = log(all.res[i, "test.per"] / (1 - all.res[i, "test.per"])) - log(all.res[i, "control.per"] / (1 - all.res[i, "control.per"])),
  #"z.age" = 0,
  #"sexm" = 0,
  "z.trial" = 0
  #"z.session" = 0
  )

  LP <- m.mat[, names(coefs)] %*% coefs # LP wrt fixed effects

  # add random effect to linear predictor:
  LP <- LP + rnorm(n = n.subject, sd = all.res[i, "r.effect"])[as.numeric(xdata$subj.id)] #+
  #+  rnorm(n = n.subject, sd = all.res[i, "r.slope.int"])[as.numeric(xdata$subj.id)] * xdata$condition.dummy +
  #+  rnorm(n = n.subject, sd = all.res[i, "r.slope.trial"])[as.numeric(xdata$subj.id)] * xdata$z.trial +
  #+  rnorm(n = n.subject, sd = all.res[i, "r.slope.session"])[as.numeric(xdata$subj.id)] * xdata$z.session 

  # generate response:
  xdata$correct <- rbinom(n = nrow(xdata), size = 1, prob = exp(LP) / (1 + exp(LP)))


  # fit full model:
  full <- keepWarnings(glmer(correct ~ condition + z.trial + (1 | subj.id),
    data = xdata, family = binomial, control = contr
  ))

  # store results:
  all.res[i, c("icpt", "conditiontest", "z.trial")] <- fixef(full$value) #, "z.session"
  all.res[i, "re.sd"] <- as.data.frame(summary(full$value)$varcor)[1, "sdcor"]
  all.res[i, "warns.full"] <- nchar(paste(full$warnings, collapse = ""))
  all.res[i, "lrt.p.con"] <- as.data.frame(drop1(full$value, test = "Chisq"))["condition", "Pr(Chi)"]
    all.res[i, "lrt.trial.p"] <- as.data.frame(drop1(full$value, test = "Chisq"))["z.trial", "Pr(Chi)"]
    print(i)
}

save.image("stealing_behind_tunnels_80subj.RData")
```

## Evaluation of results 

* number of warning per combinations of random effects (out of 1000 models per cell)  
Full model:  
```{r echo=FALSE}
#full model
tapply(X=all.res[, "warns.full"]>0, INDEX=all.res[, c("test.per", "r.effect")],
FUN=sum)
#warning codes: 
#363: unable to evaluate scaled gradient. Model failed to converge: degenerate  Hessian with 1 negative eigenvalues
#205: Model is nearly unidentifiable: large eigenvalue ratio - Rescale variables?
```


## Only models that converged are evaluated from here on:  

```{r include=FALSE}
all.res2=subset(all.res, warns.full==0)
```


### How many models converged, have a significant full-null model comparison, and a significant LRT of condition?  
```{r echo=FALSE}
n.converged<- all.res2%>%
    group_by(control.per,test.per, r.effect ) %>% #, r.slope.int
  summarise(n.converged=length(lrt.p.con))


lrt.data2 <- all.res2 %>%
  #filter(full.null.p<0.05)%>%
  group_by(control.per, test.per, r.effect ) %>% #, r.slope.int
  summarise(lrt.p.con.median = median(lrt.p.con), 
            lrt.p.trial.median = median(lrt.trial.p),
            n.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]),
            n.sign.lrt.trial = length(lrt.trial.p[lrt.trial.p < 0.05]),
            n.lrt = n.simus,
            proportion.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]) / n.simus,
            proportion.sign.lrt.trial = length(lrt.trial.p[lrt.trial.p < 0.05]) / n.simus)%>%
  full_join(n.converged)

lrt.data2
```

#### Plotting the proportion of significant LRTs for the predictor variable condition ONLY based on models that converged and with a significant full-null model comparison

```{r echo=FALSE}
p.con.power_80 <- ggplot(data = lrt.data2, aes(x= as.factor(r.effect),y = proportion.sign.lrt.con, fill=as.factor(test.per))) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1))+
  geom_hline(yintercept = 0.8, colour = "black", lwd = 1, lty = 2) +
   # geom_hline(yintercept = 0.05, colour = "darkgrey", lwd = 1.1, lty = 4) +
  scale_fill_manual(values=c("dodgerblue", "darkorange"))+
  labs(fill = "Test condition", y="Power", x= "Random intercept") +
  theme_few()+
  ggtitle("80 subject, 1 trial per condition")+
  geom_text(aes(label=round(proportion.sign.lrt.con,2)), position=position_dodge(width=0.9), vjust=-0.25)
  #theme(legend.position="none")
p.con.power_80

ggsave(p.con.power_80, filename = "graphics/stealing_behind_tunnels_80subj_power.png", scale = 0.7, height = 5, width = 8)
```



###not updated from here###
#### Plotting the intercepts


```{r echo=FALSE}
ggplot(data = all.res2, aes(x=as.factor(r.effect), y=icpt))+
  geom_jitter( alpha=0.5, col="grey")+
  geom_boxplot(aes(x=as.factor(r.effect), y=icpt, group=r.effect), alpha=0.1, outlier.colour="white")+
  geom_hline(yintercept=coefs["(Intercept)"], colour="red", lty=2)
```



#### Plotting the fixed effect of condition

```{r echo=FALSE}
p.con <- ggplot(data = all.res2, aes(x= as.factor(r.effect), fill=as.factor(test.per)))  +
  geom_jitter(data = all.res2, aes(x = as.factor(r.effect), y = conditiontest, color = as.factor(test.per)), size = 1.5, position = position_jitterdodge(dodge.width = 0.8, jitter.width = 0.5), alpha = .1) +
  scale_color_manual(values = c("dodgerblue", "darkorange")) +
    geom_boxplot(data = all.res2 %>% filter(test.per == "0.15"), aes(x = as.factor(r.effect), y = conditiontest), position = position_nudge(x = -.2), width = 0.3, alpha = 0.15, outlier.colour = "white") +
  geom_boxplot(data = all.res2 %>% filter(test.per == "0.25"), aes(x = as.factor(r.effect), y = conditiontest), position = position_nudge(x = .2), width = 0.3, alpha = 0.15, outlier.colour = "white") +
  geom_hline(aes(yintercept = test.per[1]), colour = "black", lwd = 1.1, lty = 2, alpha = 0.7) +
  geom_hline(aes(yintercept = qlogis(test.per[2])), colour = "darkgrey", lwd = 1.1, lty = 4, alpha = 0.7) +
  ylab("Condition (fixed effect)") +
  xlab("Random intercepts") +
  #ylim(-10, 10)+
  theme_few() #+
  #theme(legend.position = "none")

p.con 
```


```{r echo=FALSE}
p.legend <- ggplot(data = lrt.data2, aes(x= as.factor(control.per),y = proportion.sign.lrt.con, fill=as.factor(test.per))) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1))+
  geom_hline(yintercept = 0.95, colour = "black", lwd = 1, lty = 2) +
    geom_hline(yintercept = 0.05, colour = "darkgrey", lwd = 1, lty = 4) +
  scale_fill_manual(values=c("dodgerblue", "darkorange"))+
  labs(fill = "Functional condition", y="Power", x= "Intrinsic condition") +
  theme_few()+
  theme(legend.position="top")

p.leg <- get_legend(p.legend)

p.con <- plot_grid(perf.plot, p.con.power, p.con.str, p.con.intr, labels = c("A", "B", "C", "D"), rel_widths = c(1, 1, 1, 1), nrow=2)
p.con2 <- ggdraw(plot_grid(p.con, p.leg, ncol = 1, nrow = 2, rel_heights = c(1, 0.1)))

p.con2
#ggsave(p.con2, filename = "graphics/Explanation seeking_children simulation.png", scale = 0.65, height = 12, width = 12)
```
