---
title: "Dogs stealing behind tunnels: power simulation"
author: "Christoph Völter"
date: "15/09/2022"
output: 
  html_document:
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(cowplot)
library("gghalves")
library(ggthemes)

load("stealing_behind_tunnels_different_sample_sizes_multtrials.RData")
```

Notes:
DV: choice of opaque side
Design: 2 conditions (test / control), within subject design
1 session, 2 trials per condition (increase if needed)
Prediction: opaque side in 65% of test trials, and 50% in control trials
N = 80

## Generate data
```{r echo=FALSE, include=FALSE}
set.seed(1)
n.subject <- c(50, 60, 70, 80) # number subjects
n.per.subject <- 4 # observations per subject
n.per.condition <- 2 # observations per subject and condition
n.sessions <- 1
#age_range <- c(12:130) # age range between 1 and 13 years
test.per<-c(0.65, 0.7) # test condition
control.per<-c(0.5) # control condition
```

## Simulation

```{r eval=FALSE, include=FALSE}
n.simus <- 1000
r.effects <- c(0.62, 1.24) # random effects to be simulated
# with the coefficient of condition being 0.62 (qlogis(0.65)) we assume a moderately large random intercept of 0.62.

r.slope.con <- c(0.62)
# with the estimate being -1.098612 (qlogis(0.25)-qlogis(0.5)) we assume a moderately large random slope of 1.1.

#r.slope.trial <- 0.2
```


```{r eval=FALSE, include=FALSE}
all.res.all.samples <- data.frame()
# load packages needed:
library(lme4)
# Loading required package: Matrix
library(kyotil) # we want to store info about convergence issues

# define control structure to make convergence more likely:
contr <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))

for (j in 1:length(n.subject)) {

    # create object to store the simulation parameters and results:
  all.res <- data.frame(expand.grid(
    n.per.subject = n.per.subject, r.effect = r.effects,
    r.slope.con = r.slope.con, # r.slope.trial = r.slope.trial,
    #r.slope.session = r.slope.session,
    test.per = test.per,
    control.per = control.per,
    simu = 1:n.simus
  ))
  all.res$n.subject<-n.subject[j]
  all.res$icpt <- NA
  all.res$conditiontest <- NA
  all.res$re.sd <- NA
  all.res$warns.full <- NA
  all.res$warns.null <- NA
  all.res$lrt.p.con <- NA
  all.res$lrt.p.order <- NA
  all.res$lrt.p.trial <- NA
    all.res$full.null.p <- NA
  
  all.ests <- matrix(NA, nrow = n.simus, ncol = 1)
  colnames(all.ests) <- c("lrt.p.con")
  
  subj.id <- as.factor(paste("subj", str_pad(1:n.subject[j], 2, pad = "0"), sep = "."))
  start.data <- data.frame(expand.grid(subj.id = subj.id, condition = c("test", "control"), trial_within = c(1,2)))
  start.data <- arrange(start.data, subj.id, condition)
  start.data$trial <- c(rep(c(1,1,2,2),n.subject[j]/2), rep(c(2,2,1,1),n.subject[j]/2))
  start.data$z.trial <- as.vector(scale(as.numeric(start.data$trial)))
  start.data$order <- c(rep(rep("test_first", n.per.subject),n.subject[j]/2), rep(rep("control_first", n.per.subject),n.subject[j]/2))
  
  start.data$condition <- as.factor(start.data$condition)
  start.data$condition <- relevel(start.data$condition, ref = "control")
  start.data$condition.dummy <- as.numeric(start.data$condition == levels(start.data$condition)[2])

# center condition for random slopes:
start.data$condition.c <- as.numeric(start.data$condition) - mean(as.numeric(start.data$condition))
  
  start.data$order <- as.factor(start.data$order)
  start.data$order <- relevel(start.data$order, ref = "control_first")
  xdata <- start.data
  m.mat <- model.matrix(object = ~condition + order + z.trial, data = xdata) # create model martix

  # run simulation
    for (i in 1:nrow(all.res)) {
    set.seed(i) # allows to later replicate individual simulations
  
      coefs <- c(
        "(Intercept)" = qlogis(all.res[i, "control.per"]),
        "conditiontest" = log(all.res[i, "test.per"] / (1 - all.res[i, "test.per"])) - log(all.res[i, "control.per"] / (1 - all.res[i, "control.per"])),
      #"z.age" = 0,
      #"sexm" = 0,
      "z.trial" = 0,
      "ordertest_first" = 0
      #"z.session" = 0
      )
    
      LP <- m.mat[, names(coefs)] %*% coefs # LP wrt fixed effects
    
      # add random effect to linear predictor:
      LP <- LP + rnorm(n = n.subject[j], sd = all.res[i, "r.effect"])[as.numeric(xdata$subj.id)] +
      +  rnorm(n = n.subject[j], sd = all.res[i, "r.slope.con"])[as.numeric(xdata$subj.id)] * xdata$condition.c #+
      #+  rnorm(n = n.subject, sd = all.res[i, "r.slope.trial"])[as.numeric(xdata$subj.id)] * xdata$z.trial +
      #+  rnorm(n = n.subject, sd = all.res[i, "r.slope.session"])[as.numeric(xdata$subj.id)] * xdata$z.session 
    
      # generate response:
      xdata$correct <- rbinom(n = nrow(xdata), size = 1, prob = exp(LP) / (1 + exp(LP)))
    
    
      # fit full model:
      full <- keepWarnings(glmer(correct ~ condition + order + z.trial + (1 + condition.c | subj.id),
        data = xdata, family = binomial, control = contr
      ))
    
      # store results:
      all.res[i, c("icpt", "conditiontest", "ordertest_first", "z.trial")] <- fixef(full$value) #, "z.session"
      all.res[i, "re.sd"] <- as.data.frame(summary(full$value)$varcor)[1, "sdcor"]
      all.res[i, "warns.full"] <- nchar(paste(full$warnings, collapse = ""))
      all.res[i, "lrt.p.con"] <- as.data.frame(drop1(full$value, test = "Chisq"))["condition", "Pr(Chi)"]
      all.res[i, "lrt.p.order"] <- as.data.frame(drop1(full$value, test = "Chisq"))["order", "Pr(Chi)"]
        all.res[i, "lrt.p.trial"] <- as.data.frame(drop1(full$value, test = "Chisq"))["z.trial", "Pr(Chi)"]
        print(c("i=",i, " j=" ,j))
    }
  all.res.all.samples<- rbind(all.res.all.samples, all.res)
}

save.image("stealing_behind_tunnels_different_sample_sizes_multtrials.RData")
```

## Evaluation of results 

* number of warning per combinations of random effects (out of 1000 models per cell)  
Full model:  
```{r echo=FALSE}
#full model
tapply(X=all.res.all.samples[, "warns.full"]>0, INDEX=all.res.all.samples[, c("test.per", "r.effect")],
FUN=sum)
#warning codes: 
#363: unable to evaluate scaled gradient. Model failed to converge: degenerate  Hessian with 1 negative eigenvalues
#205: Model is nearly unidentifiable: large eigenvalue ratio - Rescale variables?
```


## Only models that converged are evaluated from here on:  

```{r include=FALSE}
all.res2=subset(all.res.all.samples, warns.full==0)
```


### How many models converged, have a significant full-null model comparison, and a significant LRT of condition?  
```{r echo=FALSE}
n.converged<- all.res2%>%
    group_by(n.subject, control.per,test.per, r.effect ) %>% #, r.slope.int
  summarise(n.converged=length(lrt.p.con))


lrt.data2 <- all.res2 %>%
  #filter(full.null.p<0.05)%>%
  group_by(n.subject, control.per, test.per, r.effect ) %>% #, r.slope.int
  summarise(lrt.p.con.median = median(lrt.p.con), 
            lrt.p.trial.median = median(lrt.p.trial),
            lrt.p.order.median = median(lrt.p.order),
            n.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]),
            n.sign.lrt.trial = length(lrt.p.trial[lrt.p.trial < 0.05]),
            n.sign.lrt.order = length(lrt.p.order[lrt.p.order < 0.05]),
            n.lrt = n.simus,
            proportion.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]) / n.simus,
            proportion.sign.lrt.trial = length(lrt.p.trial[lrt.p.trial < 0.05]) / n.simus,
            proportion.sign.lrt.order = length(lrt.p.order[lrt.p.order < 0.05]) / n.simus)%>%
  full_join(n.converged)

lrt.data2
```

#### Plotting the proportion of significant LRTs for the predictor variable condition ONLY based on models that converged and with a significant full-null model comparison
```{r echo=FALSE}
p.con.power <- ggplot(data = lrt.data2, aes(x= as.factor(r.effect),y = proportion.sign.lrt.con, fill=as.factor(test.per))) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1))+
  geom_hline(yintercept = 0.8, colour = "black", lwd = 1, lty = 2) +
   # geom_hline(yintercept = 0.05, colour = "darkgrey", lwd = 1.1, lty = 4) +
  scale_fill_manual(values=c("dodgerblue", "darkorange"))+
  labs(fill = "Test condition", y="Power", x= "Random intercept") +
  theme_few()+
  ggtitle("Different sample sizes (50 - 80), 2 trial per condition")+
  geom_text(aes(label=round(proportion.sign.lrt.con,2)), position=position_dodge(width=0.9), vjust=-0.25)+
  facet_wrap(~n.subject)
  #theme(legend.position="none")
p.con.power

ggsave(p.con.power, filename = "graphics/stealing_behind_tunnels_4_trials_differentNs_power.png", scale = 0.55, height = 12, width = 16)
```




####not updated from here
```{r}
pg<-plot_grid(p.con.power_4tr, p.con.power_8tr, rel_widths = c(1,1.5))
ggsave(pg, filename = "graphics/kea_explanation_seeking_power_plot_grid.png", width = 12, height=5, scale=0.75)
```



#### Plotting the intercepts


```{r echo=FALSE}
ggplot(data = all.res2, aes(x=as.factor(r.effect), y=icpt))+
  geom_jitter( alpha=0.5, col="grey")+
  geom_boxplot(aes(x=as.factor(r.effect), y=icpt, group=r.effect), alpha=0.1, outlier.colour="white")+
  geom_hline(yintercept=coefs["(Intercept)"], colour="red", lty=2)
```



#### Plotting the fixed effect of condition

```{r echo=FALSE}
p.con <- ggplot(data = all.res2, aes(x= as.factor(control.per), fill=as.factor(test.per)))  +
  geom_jitter(data = all.res2, aes(x = as.factor(control.per), y = conditionint, color = as.factor(test.per)), size = 1.5, position = position_jitterdodge(dodge.width = 0.8, jitter.width = 0.5), alpha = .1) +
  scale_color_manual(values = c("dodgerblue", "darkorange")) +
    geom_boxplot(data = all.res2 %>% filter(test.per == "0.15"), aes(x = as.factor(control.per), y = conditionint), position = position_nudge(x = -.2), width = 0.3, alpha = 0.15, outlier.colour = "white") +
  geom_boxplot(data = all.res2 %>% filter(test.per == "0.25"), aes(x = as.factor(control.per), y = conditionint), position = position_nudge(x = .2), width = 0.3, alpha = 0.15, outlier.colour = "white") +
  geom_hline(data = data.frame(control.per = "0.5"), aes(yintercept = coefs["conditionint"]), colour = "black", lwd = 1.1, lty = 2, alpha = 0.7) +
  geom_hline(data = data.frame(control.per = "0.4"), aes(yintercept = qlogis(0.4) - qlogis(0.15)), colour = "darkgrey", lwd = 1.1, lty = 4, alpha = 0.7) +
  ylab("Condition (fixed effect)") +
  xlab("Intrinsic condition") +
  ylim(-10, 10)+
  theme_few() #+
  #theme(legend.position = "none")

p.con 
```


```{r echo=FALSE}
p.legend <- ggplot(data = lrt.data2, aes(x= as.factor(control.per),y = proportion.sign.lrt.con, fill=as.factor(test.per))) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1))+
  geom_hline(yintercept = 0.95, colour = "black", lwd = 1, lty = 2) +
    geom_hline(yintercept = 0.05, colour = "darkgrey", lwd = 1, lty = 4) +
  scale_fill_manual(values=c("dodgerblue", "darkorange"))+
  labs(fill = "Functional condition", y="Power", x= "Intrinsic condition") +
  theme_few()+
  theme(legend.position="top")

p.leg <- get_legend(p.legend)

p.con <- plot_grid(perf.plot, p.con.power, p.con.str, p.con.intr, labels = c("A", "B", "C", "D"), rel_widths = c(1, 1, 1, 1), nrow=2)
p.con2 <- ggdraw(plot_grid(p.con, p.leg, ncol = 1, nrow = 2, rel_heights = c(1, 0.1)))

p.con2
ggsave(p.con2, filename = "graphics/Explanation seeking_children simulation.png", scale = 0.65, height = 12, width = 12)
```
